---
title: "Prevalence estimation of percentage of women 15-49 classified as having any anemia, by senatorial zones in Nigeria, using the surveyprev package"
author: "John Omoluabi" 
---

## import required libraries

```{r}
library(shapefiles)
library(surveyPrev)
library(dplyr)
library(labelled)
library(kableExtra)
library(sf)
library(sn)
library(ggplot2)
library(tidyverse)  # most variable creation here uses tidyverse 
library(tidyselect) # used to select variables in FP_EVENTS.R
library(haven)      # used for Haven labeled DHS variables
library(labelled)   # used for Haven labeled variable creation
library(expss)    # for creating tables with Haven labeled data
library(xlsx)     # for exporting to excel
library(naniar)   # to use replace_with_na function
library(here)       # to get R project path
library(sjlabelled) # to set variables label
library(survey)  # to calculate weighted ratio for GAR
library(srvyr)
library(oce)
```

## Prevalence estimation using the direct estimation model

### import state level administrative boundary (admin 1)

```{r}
admin1.shapefile<-read_sf("boundary_states_export/boundary_states_export.shp")

```

### import the our nigeria senatorial zones shapefile

```{r}
# sz.shapefile<-read_sf("nga_lga_senatorial/Nigeria_Senatorial Zone_ADJUSTED.shp")
sz.shapefile<-read_sf("boundary_lga_senatorial_zone_edited/Nigeria_Senatorial Zone.shp")
```

### uniformity check:

```{r}
admin1.shapefile$StateName[admin1.shapefile$StateName=="Federal Capital Territory"]<-"fct"
admin1.shapefile$StateName[admin1.shapefile$StateName=="NASARAWA"]<-"NASSARAWA"
admin1.shapefile$StateName<-toupper(admin1.shapefile$StateName)
sz.shapefile$Senatorial[sz.shapefile$Senatorial=="FEDERAL CAPITAL TERRITORY"]<-"FCT"

```

### uniformity check:

```{r}
test.a<-table(rep(admin1.shapefile$StateName))
test.b<-table(rep(sz.shapefile$`State name`))

test.a<-data.frame(test.a)
test.b<-data.frame(test.b)

test.a<-test.a[order(test.a$Var1, decreasing=FALSE),]
test.b<-test.b[order(test.b$Var1, decreasing=FALSE),]

print(data.frame(admin1=test.a$Var1,sz=test.b$Var1))

levels(test.a$Var1)==levels(test.b$Var1)
```

They appear to be uniform

## import cluster gps locations shapefile

```{r}
cluster.gps.shapefile <-read_sf("cluster_gps/geo_check.shp")
cluster.gps.shapefile$ADM1NAME[cluster.gps.shapefile$ADM1NAME=="FCT ABUJA"]<-"FCT"
cluster.gps.shapefile$ADM1NAME[cluster.gps.shapefile$ADM1NAME=="NASARAWA"]<-"NASSARAWA"

```

## generate cluster information, taking cluster gps, admin 1 boundary and senatorial zone boundary as parameters

```{r}
sz.cluster.info <- clusterInfo(geo=cluster.gps.shapefile, 
                               poly.adm1=admin1.shapefile, 
                               poly.adm2=sz.shapefile,
                               by.adm1 = "StateName", # join by "StateName" and "NAME_1_5" from two boundary shapefiles
                               by.adm2 = "Senatorial")
```

## function to check for the correctness of the cluster info exercise.

```{r}
check_join<-function(i){
  input<-str_split(i,"_")
  x<-grepl(input[[1]][1],input[[1]][2],fixed=TRUE)
  if(x!=TRUE){
    TRUE
    }
  else{
    FALSE
  }
  
}
sz.cluster.info$data$check_err<-map(sz.cluster.info$data$admin2.name.full,check_join)
sz.cluster.info$data$check_err<-unlist(sz.cluster.info$data$check_err)

```

```{r}
E<-sz.cluster.info$data[sz.cluster.info$data$check_err==TRUE,]
E
sf::st_write(E,"outputs/E.shp")

E<-read_sf("outputs/E.shp")
E
error.clusters<-E$cluster
error.clusters
```

```{r}
sz.cluster.info$data[sz.cluster.info$data$check_err==TRUE,]
```

*the above must amount to an empty dataframe to be correct. means that there are no errors.*

## run direct estimation model

```{r}
result <- directEST(data = data,
                       cluster.info =sz.cluster.info,
                       admin = 2,
                       aggregation = FALSE
                       )
```

```{r}
print(result$res.admin2)
```

visualize the direct estimate coefficient of variation

```{r}
result$res.admin2$Senatorial<-result$res.admin2$admin2.name
res.gps<-inner_join(result$res.admin2,sz.shapefile,by=NULL,) #join the result with cluster gps
sf::st_write(res.gps,"outputs/result_4-12-24.shp")
res.gps<-read_sf("outputs/result_4-12-24.shp")
plot(res.gps[,c("drct_st","cv")])
```

```{r}
result_nar$res.admin2$Senatorial<-result_nar$res.admin2$admin2.name
res_nar.gps<-inner_join(result_nar$res.admin2,sz.shapefile,by=NULL,)
sf::st_write(res_nar.gps,"outputs/result_nar_4-12-24.shp")
res_nar.gps<-read_sf("outputs/result_nar_4-12-24.shp")
plot(res_nar.gps[,c("drct_st","cv")])
```

```{r}
res_nar.gps
```

## CUSTOM INDICATOR (ATTENDING PRIMARY SCHOOL)

#### get dhs data needed for out indicator, Houshold recode, births recode and household member recode

```{r}
year <- 2018
country <- "Nigeria"
nar.gar.Data <- getDHSdata(country = country, 
                      indicator = NULL,
                      Recode = c("Household Member Recode","Births Recode","Household Recode"), 
                      year = year)
```

### FUNCTION DEFINITION

```{r}

######################################################
# for 2018 survey, school starts in september (9th month)
school_start_yr = 2018
school_start_mo = 9
# also we need the age ranges for primary and secondary
# for example, for Nigeria 2018, the age range is 6-11 for primary school and 13-18 for secondary school
age_prim_min = 6
age_prim_max = 11
age_sec_min = 13
age_sec_max = 18
HRdata<-nar.gar.Data$`Household Recode`
binary.indicator.gen <- function(BRdata,PRdata) {
  temp <- BRdata
  
  # keep only the variables we need
  temp <- (BRdata[, c("v001", "v002","v003", "b3", "b16")])
  
  # drop if the child in the birth history was not in the household or not alive 
  temp <- temp %>%
    filter(b16!=0 & !is.na(b16))
  
  # rename key variables for matching 
  temp[["hvidx"]] <- temp[["b16"]]  
  temp[["hv001"]] <- temp[["v001"]]
  temp[["hv002"]] <- temp[["v002"]]
  temp <- (temp[, c("hvidx", "hv001", "hv002", "b3", "b16")])
  
  # sort on key variables 
  temp <- temp[order(temp$hv001, temp$hv002, temp$hvidx),]
  
  # if there are some duplicates of line number in household questionnaire, we need to drop the duplicates
  temp <- temp[!duplicated(temp[,c("hvidx", "hv001", "hv002")]),]
  
  # use the PR file for household members for the NAR and GAR indicators
  BR_PMRdata <- merge(PRdata, temp, by = c("hv001", "hv002", "hvidx"), all.y = FALSE, all.x = TRUE)
  
  #restrict to de facto household members age 5-24, and drop all others
  BR_PMRdata <- BR_PMRdata %>%
    filter(hv103==1 & inrange(hv105,5,24))
  
  
  # produce century month code of start of school year for each state and phase
  cmcSch = (school_start_yr - 1900)*12 + school_start_mo
  
  # calculate the age at the start of the school year, using the date of birth from the birth history if we have it
  # Impute an age at the beginning of the school year when CMC of birth is unknown
  # the random imputation below means that we won't get a perfect match with the report, but it will be close
  BR_PMRdata[["xtemp"]] <- ifelse(is.na(BR_PMRdata[["b3"]]), BR_PMRdata[["hv008"]] - (BR_PMRdata[["hv105"]]*12), 0)
  BR_PMRdata[["cmctemp"]] <- ifelse(is.na(BR_PMRdata[["b3"]]), BR_PMRdata[["xtemp"]] - as.integer(runif(1, 0, 1)*12), 0)
  BR_PMRdata[["school_age"]] <- ifelse(is.na(BR_PMRdata[["b3"]]), 
                                       as.integer((cmcSch - BR_PMRdata[["cmctemp"]])/12), 
                                       as.integer((cmcSch - BR_PMRdata[["b3"]])/12))
  
  # Generate variables for whether the child is in the age group for primary or secondary school
  BR_PMRdata[["prim_age"]] <- ifelse(inrange(BR_PMRdata[["school_age"]],age_prim_min,age_prim_max), 1, 0)
  BR_PMRdata[["sec_age"]] <- ifelse(inrange(BR_PMRdata[["school_age"]],age_sec_min,age_sec_max),1, 0)
  
  # create the school attendance variables, not restricted by age
  BR_PMRdata[["prim"]] <- ifelse(BR_PMRdata[["hv122"]]==1, 1, 0)
  BR_PMRdata[["sec"]] <- ifelse(BR_PMRdata[["hv122"]]==2, 1, 0)
  
  # set sample weight
  BR_PMRdata[["wt"]] = BR_PMRdata[["hv005"]]/1000000
  
  # For NAR we can use this as just regular variables and can tabulate as follows, but can't do this for GAR as the numerator is not a subset of the denominator
  # NAR is just the proportion attending primary/secondary school of children in the correct age range, for de facto children 
  BR_PMRdata[["nar_prim"]] <- ifelse(BR_PMRdata[["prim_age"]]==1, BR_PMRdata[["prim"]], 0)
  BR_PMRdata[["nar_sec"]] <- ifelse(BR_PMRdata[["sec_age"]]==1, BR_PMRdata[["sec"]], 0)
  
  BR_PMRdata <- BR_PMRdata %>%
    mutate(nar_prim = set_label(nar_prim, label = "Primary school net attendance ratio (NAR)"),
           nar_sec = set_label(nar_sec, label = "Secondary school net attendance ratio (NAR)"))

  return(BR_PMRdata)
}

```

#### generate indicator

```{r}
pre.data<-binary.indicator.gen(BRdata=BRdata,PRdata=PRdata)
```

#### create our dud function, apparently we cannot do without it

```{r}
make.value <- function(input) {
    # set_variable_labels(ph_sch_nar_prim = "Primary school net attendance ration (nar)")
  # colnames(BR_PMRdata)[colnames(BR_PMRdata) == "nar_prim"] <- "value"
  return(input)
}
```

## Run direct Estimates

```{r}

# this is redundant
pre.data$value<-0

pre.data$value[(pre.data$prim_age==1)&(pre.data$prim==1)]<-1

pre.data.prim<-pre.data %>% filter(prim_age==1)

# sz.data <- getDHSindicator(pre.data, indicator = NULL,FUN=make.value)
sz.data <- getDHSindicator(pre.data.prim, indicator = NULL,FUN=make.value) # denominator is all 6-11 year olds



# result.sz <- directEST(data = data.11,
#                        cluster.info =sz.cluster.info,
#                        admin = 2,
#                        aggregation = FALSE
# )
result.sz <- directEST(data = sz.data,
                       cluster.info =sz.cluster.info,
                       admin = 2,
                       aggregation = FALSE
)


```

## Visualize senatorial zone results

```{r}
result.sz$res.admin2$Senatorial<-result.sz$res.admin2$admin2.name
res.gps<-inner_join(result.sz$res.admin2,sz.shapefile,by=NULL,) #join the result with cluster gps
sf::st_write(res.gps,"outputs/result_11-12-24.shp")
res.gps<-read_sf("outputs/result_11-12-24.shp")
plot(res.gps[,c("drct_st","cv")])
```
